---
title: "Project 9.3 R Markdown Presentation & Plotly"
author: "ElenaKush"
date: '2020-04-26'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r environment, include=FALSE, warning = FALSE, message = FALSE}
if("pacman" %in% rownames(installed.packages()) == FALSE){ install.packages("pacman") }
library(pacman)
p_load(dplyr,
       stringr,
       tidyr,
       plotly)
```

```{r datasets, include = FALSE, warning = FALSE, message = FALSE}
local_path <- "C:/Users/kushnarevaei/Documents/GitHub/JH_R_Course/Course9/Week3/Project9.3/developing_data_products_plotly/"
URL_cases_global <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"

download.file(URL_cases_global,
              destfile = paste0(local_path,"data/cases_global.csv"))

cases_global <- read.csv(paste0(local_path,"data/cases_global.csv"))

URL_deaths_global <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"

download.file(URL_deaths_global,
              destfile = paste0(local_path,"data/deaths_global.csv"))

deaths_global <- read.csv(paste0(local_path,"data/deaths_global.csv"))

```

```{r EDA and preprocessing, include = FALSE, warning = FALSE, message = FALSE}
##head(cases_global)
##View(cases_global)
##str(cases_global)
##dim(cases_global)
##names(cases_global)
names(cases_global) <- str_remove(names(cases_global), "^X")
names(deaths_global) <- str_remove(names(deaths_global), "^X")

lastdate <- names(cases_global)[ncol(cases_global)]
lastdate <- as.Date(lastdate, "%m.%e.%y")

top15 <- cases_global %>%
    mutate(total = cases_global[, ncol(cases_global)]) %>%
    group_by(Country.Region = cases_global$Country.Region) %>%
    summarise(total = sum(total)) %>%
    arrange(desc(total)) %>%
    head(15)

cases_global <- cases_global %>%
    pivot_longer(-c(1:4), names_to = "Date", values_to = "Cases") %>%
    mutate(Date = as.Date(Date, "%m.%e.%y")) %>%
    group_by(Country.Region, Date) %>%
    summarise(Cases = sum(Cases)) %>%
    filter(Cases != 0, Country.Region %in% top15$Country.Region)

##dim(deaths_global)
##names(deaths_global)
deaths_global <- deaths_global %>%
    pivot_longer(-c(1:4), names_to = "Date", values_to = "Deaths") %>%
    mutate(Date = as.Date(Date, "%m.%e.%y")) %>%
    group_by(Country.Region, Date) %>%
    summarise(Deaths = sum(-Deaths)) %>%
    filter(Country.Region %in% top15$Country.Region)

test <- left_join(cases_global, deaths_global, by = c("Country.Region" = "Country.Region", "Date" = "Date")) %>%
    arrange(Country.Region, Date) %>%
    group_by(Country.Region) %>% 
    mutate(Days.Since.First.Case = row_number())

```

## PROJECT DESCRIPTION

For the project purposes we analysed how the COVID-19 confirmed cases and deaths were developed in different countries (top-15) since the first COVID-19 case was confirmed in each country. 

We used data posted by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University:
https://github.com/CSSEGISandData/COVID-19. 

The data is up to `r lastdate`. 

The project github repository with .Rmd file:
https://github.com/ElenaKush/developing_data_products_plotly. 

## PLOTS

```{r plot, warning = FALSE, message = FALSE}
x <- list(title = "Days since the first case in the country (for China since 2020-01-22)")

plot_cases <- plot_ly(test, x = ~Days.Since.First.Case, y = ~Cases, color = ~Country.Region, type = 'scatter', mode = 'lines', colors = "Paired", legendgroup = ~Country.Region)

plot_deaths <- plot_ly(test, x = ~Days.Since.First.Case, y = ~Deaths, color = ~Country.Region, type = 'scatter', mode = 'lines', colors = "Paired", legendgroup = ~Country.Region, showlegend = FALSE) %>%
    layout(xaxis = x, legend = list(font = list(size = 12)))

subplot(plot_cases, plot_deaths, nrows = 2, shareX = TRUE, titleY = TRUE) %>%
    layout(xaxis = x,
           annotations = list(
               list(x = 0.5 , y = 1, text = "DETECTED CASES", showarrow = F, xref='paper', yref='paper', font = list(size = 18)),
               list(x = 0.5 , y = 0.5, text = "DEATHS", showarrow = F, xref='paper', yref='paper', font = list(size = 18))
               )
           )
```

## THANK YOU